Hunting experience shapes individual foraging specialization and predator-prey interactions in an online videogame

### 2023-04-26 manuscript/Chapter2-manuscript.Rmd
update tables in all .Rmd files

Keywords: foraging behaviour, arms race, reciprocal behavioural plasticity, learning, antipredator behaviour, virtual ecology

@@

@@

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [@Griffen.etal2012;@Michalko.Pekar2016;@Moran.etal2017;@Michalko.etal2021]. Indeed, predator populations often comprise assemblages of individuals specializing in different foraging tactics or on different resources irrespective of sexual, morphological, or age-related differences [@Estes.etal2003;@Tinker.etal2008;@Kernaleguen.etal2015;@Phillips.etal2017]. A growing body of evidence suggests that ecological interactions such as predator-prey interactions can drive this individual foraging specialization [@Araujo.etal2011;@toscanoPersonalityForagingBehavior2016]. When they hunt, predators often use techniques that are fine-tuned to the type of prey that they encounter [@Davoren.etal2003;@Estes.etal2003;@Woo.etal2008;@Courbin.etal2018], and their capacity to use them effectively is contingent on periods of extensive practice (i.e. experience). While hunting experience may be essential to maintain or improve foraging success, we have few empirical assessments of its role in predator foraging specialization [@Dukas2019] and its consequences on predator-prey interactions.

@@

There is currently a lack of consensus on the fitness advantages of specialized vs flexible foraging [@Phillips.etal2017]. This is reflected in the literature showing contrasting results in the links between specialization and fitness. For instance, some studies report increasing benefits of specialization [@Pintor.etal2014;@Patrick.Weimerskirch2014a;@vandenBosch.etal2019], some report that flexible foraging has greater benefits [@Paull.etal2012;@Manlick.etal2021], and others find equal benefits depending on timescales [@Woo.etal2008;@Potier.etal2015]. This may in part be due to the limited information we have on the ecological contexts favouring specialization over flexibility in nature. However, in predator-prey systems, many studies have outlined that fluctuations in the predictability of resources throughout a predator's lifetime may be a key factor shaping foraging specialization [@Weimerskirch2007;@Woo.etal2008;@Chang.etal2017;@Phillips.etal2017;@Courbin.etal2018]. The resource-predictability (or resource-heterogeneity) hypothesis argues that when resources are predictable (or stable), individual specialists should have higher capture rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should benefit when resources fluctuate, as fine adjustments to resources are key for a predator's success and survival [@Karkarey.etal2017;@Holm.etal2019;@Santoro.etal2019]. When predator-prey interactions are directly observed, empirical evidence shows that predators often match, among other types of traits [see @Kishida.etal2006;@Hanifin.etal2008;@Brousseau.etal2018;@Reimche.etal2020], their locomotor and behavioural traits to those of their prey [@McGhee.etal2013a;@Bro-Jorgensen2013;@Chang.etal2017;@Szopa-Comley.Ioannou2022]. Yet, a recurring question that emerges from these studies is whether and how the behavioural adjustments of predators reflect individual differences in experience/learning [@Kelley.Magurran2011]. Thus, uncovering the role of learning in predator-prey systems where interactions are directly monitored would enable researchers to better predict the behavioural decisions and the success of predators when they are hunting.

The integration of individual behavioural variation in the study of predator-prey interactions has gained traction in recent years, with empirical studies revealing important consequences for habitat use, functional responses, prey choice, and foraging rate [@Kobler.etal2009;@toscanoTraitmediatedFunctionalResponses2014;@Patrick.Weimerskirch2014;@Matsumura.Miyatake2022a]. However, an important and recurring challenge impeding research on predator-prey behavioural interactions, at the individual level, is the need to collect data simultaneously on both the predator and prey. We recently demonstrated with behavioural data from an online predator-prey videogame, called *Dead by Daylight* (*DBD*), that virtual systems can overcome this challenge and help uncover the mechanisms that shape predator-prey interactions [@FraserFranco.etal2022]. For instance, we found that some individual predators hunt at high speeds, while others prefer to stalk and ambush their prey. Individuals also flexibly switch between these tactics, and their success is shaped by the speed of the prey, where faster prey are more difficult to capture. The prey need to forage for resources while paying attention to the predator to avoid being detected and chased. The prey can contribute to the group's success by healing or helping others escape the predator, or be selfish and attempt to escape by themselves [@Cere.etal2021]. Predators must learn how the prey behave and decide how best to capture them. Thus, *DBD* simulates a highly dynamic system where both predators and prey must adjust to each other, suggesting that virtual systems could be useful in identifying general ecological patterns, which might in turn help to advance the current gaps in predator-prey research. Other studies on virtual predator-prey systems show that predation regimes can drive individual variation in risk perception [@Beauchamp2020], that familiarity between prey has a positive indirect effect on survival [@Cere.etal2021], and that prey face contrasting natural and social selection regimes (Santostefano et al. in prep). Virtual predator-prey systems generate large volumes of data on interacting players throughout their lifetime in the game under realistic ecological scenarios. Hence, they offer the opportunity to tackle fundamental questions about the role of experience and prey behaviour on individual predator foraging specialization and performance.

@@

@@

# Open the loo table
path <- file.path(here::here("outputs", "03_outputs_model-validation"))
loo_tab <- data.table(readRDS(file.path(path, "03A_lootable.rds")))

@@

# Custom theme
my_theme <- function(x, ...) {
   x <- colformat_double(x, big.mark = " ",
                         decimal.mark = ".",
                         digits = 2)
   x <- set_table_properties(x, layout = "fixed")
   x <- border_remove(x)
   std_border <- fp_border(width = 1, color = "black")
   x <- hline_top(x, part = "all", border = std_border)
   x <- hline_bottom(x, part = "all", border = std_border)
   autofit(x)
}

@@

tab1 <- loo_tab %>%
   select(model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
   flextable(col_keys = my_header$col_keys) %>%
   set_header_df(mapping = my_header, key = "col_keys") %>%
   my_theme() %>%
   merge_v(part = "header") %>%
   merge_h(part = "header") %>%
   fontsize(size = 10, part = "all") %>%
   font(fontname = "Times New Roman", part = "all") %>%
   align(align = "left", part = "all", j = 1) %>%
   align(align = "center", part = "all", j = 2) %>%
   align(align = "center", part = "all", j = 3) %>%
   align(align = "center", part = "all", j = 4) %>%
   align(align = "center", part = "all", j = 5)

# ===========================================================================
# 3. Customize and print the table
# ===========================================================================

# Autofit to window
tab1 <- autofit(tab1)

# Add a footnote
tab1 <- tab1 %>%
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      "* 'elpd' refers to the expected log pointwise density and is the value chosen to select the best model."),
    ref_symbols = " "
  ) %>%
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 11, part = "all")

# Table caption
set_caption(
  tab1,
  caption = "Leave-one-out cross-validation table of the three hierarchical GAMMs relating hunting success to player experience",
  style = "Table Caption"
)

@@

```{r fig1, out.width = "100%", fig.cap = "**Figure 1.** Among individual differences in the development of hunting expertise. The predators' hunting success (i.e. the probability of capturing the four prey) is on the y axis, and the predators' cumulative experience (i.e. the number of matches played prior to each observation) is on the x axis. Each fitted curve represents an individual predator. (A) A generalized additive mixed model where we do not control for prey speed (B) A generalized additive mixed model where we control for prey speed."}

knitr::include_graphics(
  here::here("outputs", "05_outputs_figures", "05_figure1.png")
)

At the population level, predators did not change their mean speed with experience (i.e. intercept of the mean predator speed; Table 2). Moreover, there was no directional increase in foraging flexibility or specialization with experience. Indeed, the observed pattern is more complex given that predators became more flexible with experience at intermediate levels of experience (i.e. intercept of sigma for predator speed is larger for intermediates than novices; Table 2 and Appendix S1: Figure S4), but then return to novice-levels of flexibility with advanced experience (i.e. intercept of sigma for predator speed is smaller for advanced players than intermediates; Table 2 and Appendix S1: Figure S4). Thus, our analyses suggest that population-level behavioural specialization changes nonlinearly (quadratic) with experience rather than linearly.

@@

# Import the model
fit <- readRDS(
  file.path(path, "02_outputs_models", "02B_DHMLM.rds")
)

@@

  # table structure
  table %>%
  select(trait, parameter,
         novice, intermediate, advanced) %>%
  flextable(col_keys = my_header$col_keys) %>%
  set_header_df(mapping = my_header, key = "col_keys") %>%
  my_theme() %>%
  # Align repeated rows on column 1
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top", part = "body") %>%
  # font
  fontsize(size = 10, part = "all") %>%
  font(fontname = "Times New Roman", part = "all") %>%

  # Align text
  align(align = "left", part = "all", j = 1) %>%
  align(align = "left", part = "all", j = 2) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5)

# ===========================================================================
# 4. Customize and print the table
# ===========================================================================

# Autofit to window size
tab2 <- autofit(tab2)

# Table footnote
tab2 <- tab2 %>%
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "We exponentiated the dispersion parameters (i.e. sigma) which are estimated on a log scale. We back-transformed the hunting success values, estimated on a logit scale, back to a probability scale.",
      as_sup("\nb "),
      "The intercept values on the mean part of the equation for all traits indicate mean behaviour and success at the population level. The intercept values on the dispersion (i.e. sigma) part of the equation for predator speed indicate behavioural specialization at the population level."
    ),
    ref_symbols = " "
  ) %>%
  font(fontname = "Times New Roman", part = "all") %>%
  width(j = c(2, 3, 4, 5), width = 3.7, unit = "cm") %>%
  width(j = 1, width = 3.2, unit = "cm")

# Table caption
set_caption(
  tab2,
  caption = "Posterior means and 95% credible intervals of the fixed effects estimated by the MDHGLM of predator speed, prey speed, and predator hunting success.",
  style = "Table Caption"
)

@@

At the individual level, we detected differences in the mean movement speed among predators (i.e. the intercept of the mean speed for the predator ID; Table 3). These differences remained stable from the novice to the intermediate stages, and then increased at the advanced stage to reach a level higher than when predators were novice (Table 3 and Appendix S1: Figure S4). Predators displayed considerable differences in their degree of hunting specialization, and these differences increased with experience (i.e. the intercept of sigma for the predator ID increases with experience; Table 3 and Appendix S1: Figure S4). Thus, as individual predators gain experience, the population becomes more heterogeneous, with an assemblage of both specialist and flexible hunters. Moreover, some individual predators switched from a flexible to a specialized hunting tactic, and vice-versa, as they gained experience (Figure 2). Importantly, 12.6% of the predator population displayed a large change in foraging specialization/flexibility with experience (i.e. $\ge$0.2 change in standard deviation), 44.7% displayed lower changes (i.e. >0.05 and <0.2 change in standard deviation), while 42.7% remained relatively stable (i.e. $\le$0.05 change in standard deviation).

@@

# Custom theme
my_theme <- function(x, ...) {
  x <- colformat_double(x, big.mark = " ",
                        decimal.mark = ".",
                        digits = 3)
  x <- set_table_properties(x, layout = "fixed")
  x <- border_remove(x)
  std_border <- fp_border(width = 1, color = "black")
  x <- hline_top(x, part = "all", border = std_border)
  x <- hline_bottom(x, part = "all", border = std_border)
  autofit(x)
}

@@

  # Align text
  align(align = "left", part = "all", j = 1) %>%
  align(align = "left", part = "all", j = 2) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5)

# ===========================================================================
# 4. Customize and print the table
# ===========================================================================

# Autofit to window
tab3 <- autofit(tab3)

# Add footnote
tab3 <- tab3 %>%
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "We exponentiated the dispersion parameters (i.e. sigma) which are estimated on a log scale. All the reported values are standard deviations.",
      as_sup("\nb "),
      "The intercept values on the mean part of the equation for all traits indicate among individual differences in mean behaviour and success.",
      as_sup("\nc "),
      "The intercept values on the dispersion (i.e. sigma) part of the equation for predator speed indicate among individual differences in behavioural specialization. For prey speed and hunting success, they indicate among individual differences in the variability of prey encounters and variability in hunting success, respectively."
    ),
    ref_symbols = " "
  ) %>%
  font(fontname = "Times New Roman", part = "all") %>%
  width(j = c(2, 3, 4, 5), width = 3.7, unit = "cm") %>%
  width(j = 1, width = 3.2, unit = "cm")

# Add table caption
set_caption(
  tab3,
  caption = "Posterior means and 95% credible intervals of the random effects estimated by the MDHGLM of predator speed, prey speed, and predator hunting success.",
  style = "Table Caption"
)

@@

As we predicted, predators differed in the mean speed of the prey that they encountered across matches. Thus, some individuals encountered prey that were on average slower, while others encountered prey that where on average faster throughout the study period (i.e. the intercept of the mean prey speed for the predator ID; Table 3). These differences were greater when predators reached an advanced level of hunting experience (i.e. the intercept of the mean prey speed for the predator ID was higher at the advanced stage; Table 3 and Appendix S1: Figure S4). In addition, individuals differed in the heterogeneity of the groups of prey encountered across the study period. Hence, some predators encountered prey groups with similar average speeds, while others encountered prey groups with different average speeds throughout the study period (i.e. the intercept of sigma prey speed for the predator ID; Table 3).

```{r fig2, out.width = "100%", fig.cap = "**Figure 2.** Individual behavioural distribution of the predators' speed drawn from the MDHGLM estimates for individuals that had (A) the greatest increase in specialization and (B) the greatest increase in flexibility. The background (gray) distributions are for when individuals were novice hunters, and the foreground (blue) distributions with solid contour lines are for when they were advanced hunters. Both figure panels are ordered by ascending degree of increase in either specialization or flexibility. We selected the individuals displayed in the figure by first substracting the estimated standard deviation of all individuals as novices with their standard deviation as advanced hunters, and then selecting those with the greatest difference using the 25% and 75% quantiles. Individuals on panel A have an increase in specialization equal or greater than 0.2 standard deviations, while those on panel B have an increase in flexibility equal or greater than 0.28 standard deviations."}

knitr::include_graphics(
  here::here("outputs", "05_outputs_figures", "05_figure2.png")
)

>>>>

Keywords: foraging behaviour, reciprocal behavioural plasticity, learning, antipredator behaviour, virtual ecology

@@

Many predator-prey systems are driven by reciprocal interactions where one responds to changes in the behaviour of the other. However, the mechanisms driving these individual responses are only beginning to be explored, in part because of the challenges associated with monitoring direct predator-prey interactions. In this study, we used a virtual predator-prey system (the online game *Dead by Daylight*) to test how hunting experience shapes individual foraging specialization and success in predators across repeated interactions with their prey. With increasing experience, predators fine-tuned their tactics to the behaviour of their prey. Individuals that adopted a flexible tactic (i.e. generalists) encountered groups of prey with varying speeds, and those that adopted a cursorial and specialized tactic (i.e. specialists) encountered groups of prey with similar speeds throughout the study period. Our results suggest that experience enables predators to learn how to optimize their tactics to the type of prey that they encounter (i.e. fast or slow), which may act as a stabilizing mechanism promoting behavioural diversification in predator-prey systems.

@@

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [@Griffen.etal2012;@Michalko.Pekar2016;@Moran.etal2017;@Michalko.etal2021]. Indeed, predator populations often comprise assemblages of individuals specializing in different foraging tactics or on different prey irrespective of sexual, morphological, or age-related differences [@Estes.etal2003;@Tinker.etal2008;@Kernaleguen.etal2015;@Phillips.etal2017]. A growing body of evidence suggests that reciprocal interactions between predators and prey could drive this individual foraging specialization [@Araujo.etal2011;@toscanoPersonalityForagingBehavior2016;@DeLisle.etal2022]. For instance, predators often hunt using techniques that are fine-tuned to the type of prey that they encounter [@Davoren.etal2003;@Estes.etal2003;@Woo.etal2008;@Courbin.etal2018], and their capacity to use them effectively is contingent on periods of extensive practice (i.e. experience). While hunting experience may be essential to maintain or improve foraging success, we have few empirical assessments of its role in predator foraging specialization [@Dukas2019] and its consequences on predator-prey interactions.

@@

There is currently a lack of consensus on the fitness advantages of specialized vs flexible foraging [@Phillips.etal2017]. This is reflected in the literature showing contrasting results in the links between specialization and fitness. For instance, some studies report increasing benefits of specialization [@Pintor.etal2014;@Patrick.Weimerskirch2014a;@vandenBosch.etal2019], some report that flexible foraging has greater benefits [@Paull.etal2012;@Manlick.etal2021], and others find equal benefits depending on timescales [@Woo.etal2008;@Potier.etal2015]. This may in part be due to the limited information we have on the ecological contexts favouring specialization over flexibility in nature. However, in predator-prey systems, many studies have outlined that fluctuations in the predictability of resources throughout a predator's lifetime may be a key factor shaping foraging specialization [@Weimerskirch2007;@Woo.etal2008;@Chang.etal2017;@Phillips.etal2017;@Courbin.etal2018]. The resource-predictability (or resource-heterogeneity) hypothesis argues that when resources are predictable (or stable), individual specialists should have higher capture rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should benefit when resources fluctuate, as fine adjustments to resources are key for a predator's success and survival [@Karkarey.etal2017;@Holm.etal2019;@Santoro.etal2019]. When predator-prey interactions are directly observed, empirical evidence shows that predators often match their locomotor and behavioural traits to those of their prey [@McGhee.etal2013a;@Bro-Jorgensen2013;@Chang.etal2017;@Szopa-Comley.Ioannou2022]. Yet, a recurring question emerging from these studies is whether the mechanism driving such behavioural adjustments by predators involves individual differences in experience/learning [@Kelley.Magurran2011]. Thus, uncovering the role of learning in predator-prey systems where interactions are directly monitored would enable researchers to better predict the behavioural decisions and the success of predators when they are hunting.

The integration of individual behavioural variation in the study of predator-prey interactions has gained traction in recent years, with empirical studies revealing important consequences for habitat use, functional responses, prey choice, and foraging rate [@Kobler.etal2009;@toscanoTraitmediatedFunctionalResponses2014;@Patrick.Weimerskirch2014;@Matsumura.Miyatake2022a]. However, an important and recurring challenge impeding research on predator-prey behavioural interactions, at the individual level, is the need to collect data simultaneously on both the predator and prey. We recently demonstrated, using an online predator-prey videogame called *Dead by Daylight* (*DBD*), that virtual systems can overcome this challenge and help uncover the mechanisms that shape predator-prey interactions [@FraserFranco.etal2022]. In this dynamic virtual system, four prey players need to forage for resources while avoiding a predator player. The prey can contribute to the group's success by healing or helping others escape the predator, or be selfish and attempt to escape by themselves [@Cere.etal2021]. The predators' hunting tactics and their success are largely driven by the movement of the prey, where faster prey are more difficult to capture [@FraserFranco.etal2022]. Predators should thus benefit from learning how the prey behave to be successful. Other studies using virtual predator-prey systems show that predation regimes drives individual variation in risk perception [@Beauchamp2020], that familiarity between prey has a positive indirect effect on survival [@Cere.etal2021], and that prey face contrasting natural and social selection regimes (Santostefano et al. in prep). These examples outline how virtual systems could be useful in identifying general ecological patterns, which might in turn help to advance the current gaps in predator-prey research. They also generate large volumes of data on interacting players throughout their lifetime in the game under realistic ecological scenarios. Hence, virtual systems offer the opportunity to tackle fundamental questions about the role of experience and prey behaviour on individual predator foraging specialization and performance.

@@

At the population level, predators did not change their mean speed with experience (i.e. intercept of the mean predator speed; Table 2). Moreover, there was no directional increase in foraging flexibility or specialization with experience. Indeed, the observed pattern is more complex given that predators became more flexible with experience at intermediate levels of experience (i.e. intercept of sigma for predator speed is larger for intermediates than novices; Table 2 and Appendix S1: Figure S3), but then return to novice-levels of flexibility with advanced experience (i.e. intercept of sigma for predator speed is smaller for advanced players than intermediates; Table 2 and Appendix S1: Figure S3). Thus, our analyses suggest that population-level behavioural specialization changes nonlinearly (quadratic) with experience rather than linearly.

At the individual level, we detected differences in the mean movement speed among predators (i.e. the intercept of the mean speed for the predator ID; Table 3). These differences remained stable from the novice to the intermediate stages, and then increased at the advanced stage to reach a level higher than when predators were novice (Table 3 and Appendix S1: Figure S3). Predators displayed considerable differences in their degree of hunting specialization, and these differences increased with experience (i.e. the intercept of sigma for the predator ID increases with experience; Table 3 and Appendix S1: Figure S3). Thus, as individual predators gain experience, the population becomes more heterogeneous, with an assemblage of both specialist and flexible hunters. Moreover, some individual predators switched from a flexible to a specialized hunting tactic, and vice-versa, as they gained experience (Figure 2). Importantly, 12.6% of the predator population displayed a large change in foraging specialization/flexibility with experience (i.e. $\ge$0.2 change in standard deviation), 44.7% displayed lower changes (i.e. >0.05 and <0.2 change in standard deviation), while 42.7% remained relatively stable (i.e. $\le$0.05 change in standard deviation).

As we predicted, predators differed in the mean speed of the prey that they encountered across matches. Thus, some individuals encountered prey that were on average slower, while others encountered prey that where on average faster throughout the study period (i.e. the intercept of the mean prey speed for the predator ID; Table 3). These differences were greater when predators reached an advanced level of hunting experience (i.e. the intercept of the mean prey speed for the predator ID was higher at the advanced stage; Table 3 and Appendix S1: Figure S3). In addition, individuals differed in the heterogeneity of the groups of prey encountered across the study period. Hence, some predators encountered prey groups with similar average speeds, while others encountered prey groups with different average speeds throughout the study period (i.e. the intercept of sigma prey speed for the predator ID; Table 3).

We predicted that this would occur if, by chance, predators differed in the prey that they encountered, because some may be easier and others more difficult to capture. In nature, prey can use a variety of defenses such as physical armaments, toxins, camouflage, or antipredator behaviours to escape predation [@BrodieIII.BrodieJr.1999;@Bowen.etal2002;@brodieCoEvolutionPredatorsPrey2010;@Nomura.etal2011;@Carey.Wahl2011]. In *DBD* we previously showed that faster movement is an effective strategy used by the prey to avoid predation [@FraserFranco.etal2022], and other studies have found that as well [@Walker.etal2005;@Kelley.Magurran2011;Martin.etal2022]. Yet, prey speed did not shape individual differences in expertise acquisition, but other antipredator tactics may be involved. For example, it is well known in *DBD* that some prey groups use mobbing to confuse the predator or prevent it from capturing an injured individual. However, like natural predator-prey systems, direct predator-mobbing in *DBD* is dangerous and requires coordinated actions that are learned through experience [@Graw.Manser2007;@Dutour.etal2016;@Pitman.etal2017;@Dutour.etal2019;@Szymkowiak2021], and was thus probably not a pervasive antipredator strategy. Other antipredator tactics known to confuse human and non-human predators during the attack sequence are unpredictable escapes and turning angles [@Jones.etal2011;@Briffa2013;@Herbert-Read.etal2017;@Moore.etal2017;@Richardson.etal2018]. While predators can adjust to unpredictable prey escapes [@Szopa-Comley.Ioannou2022], it could still be more arduous for a predator to learn and optimize its hunting success if it more often encountered prey using these tactics.

Alternatively, it is possible that individual variation in learning rates due to time delays between hunting events drove the observed differences in expertise acquisition, irrespective of the prey groups that predators encountered. For example, in *DBD*, a predator that played 300 matches in the span of six months might forget more critical information (e.g. prey escape patterns or muscle memory) than one that played 300 matches in the span of 6 days. While this has not been formally tested, a greater delay between exposure events is hypothesized to attenuate or even negate the relationship between success and experience [@Endler1991;@Wright.etal2022]. Moreover, the predators may have differed in their ability to detect prey. For example, it is widely demonstrated that predators learn visual cues such as coloration to recognize and avoid unpalatable prey [@Skelhorn.etal2016;@Ward-Fear.etal2020;@McLellan.etal2021], or habitat characteristics to locate food patches [@Ehlinger1989;@Hughes.etal1992;@Warburton2003]. @Szopa-Comley.etal2020 also found consistent among individual variation in the capacity to detect cryptic and conspicuous prey in three-spined sticklebacks (*Gasterosteus aculeatus*). Thus, individual predators should differ in their expertise if they do not learn equally how to exploit the visual cues that facilitate prey detection.

As predators gained experience, their tactic was increasingly matched to the speed of their prey. One interpretation could be that both predators and prey learned from each other in an arms-race-like reciprocal relationship [@Kishida.etal2006;@Kishida.etal2009;@Edgell.Rochette2009a;@Mougi.etal2011]. For instance, predators that encountered faster prey could have learned to increase their speed to better capture them, resulting in an increase in the speed of the prey to circumvent the predator's attacks. In this type of system, faster prey should be better at escaping both slower and faster predators [@Mougi.etal2011], which concords with what we previously observed in this sytem [@FraserFranco.etal2022]. @BrodieIII.BrodieJr.1999 argue that when prey express defensive phenotypes (e.g. physical armaments or venom), selection on predators should be strong because the consequences of being injured are greater than missing a meal. If predators improve their exploitative ability, then an arms-race should emerge, and the consequences of hunting dangerous prey should decrease [@BrodieIII.BrodieJr.1999]. Similarly, we found that has the predator gained experience, the negative consequences of encountering prey groups with higher speeds were relaxed, and the variation in success among predators decreased. However, in an arms-race, faster predators would have successfully preyed upon faster and slower prey and would have obtained higher success than slower predators [@Mougi.etal2011], which was not the case in the present study. A more likely scenario is that predators and prey in *DBD* were involved in a matching response relationship [@Mougi.etal2011], because faster-specialist and slower-flexible hunters preyed upon different prey, but they both achieved equal success. Thus predators in *DBD* appear to optimize their success by learning to match their tactic to their prey with experience, suggesting that learning through reciprocal interactions may attenuate selection exerted by prey on predators [@Anderson1995;@Ancel1999;@Borenstein.etal2006;@Paenke.etal2007].

Trait-mediated predator-prey interactions are important processes involved in population cycles and should occur when predators match their phenotype to their prey [@abramsEvolutionPredatorPreyInteractions2000]. Yet, it has remained largely unknown whether this trait-matching results from predators learning how to capture their prey, in part because of the challenges of investigating direct interactions in the wild. By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that individual predators displayed large differences in the development of their expertise. With experience, half of the predator population changed their behaviour by developping either a specialized or flexible tactic that was matched to the speed of their prey. Overall, both type of hunters achieved equal hunting success. Our results have potential applications for predator-prey models of trait-matching. For instance, simulation studies could incorporate individual variation in experience or learning to better predict under which ecological/evolutionary contexts specialization should be favoured over flexibility. Lastly, virtual systems are increasingly recognized among ecologists as being useful to test hypotheses on consumer-resource interactions [@Barbe.etal2020;@Beauchamp2020;@Cere.etal2021;@FraserFranco.etal2022], but also in other fields such as citizen science and conservation [@Sandbrook.etal2015;@Redpath.etal2018;@Duthie.etal2021]. We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.

We warmly thank Behaviour Interactive Inc. for our collaboration which helps in advancing predator-prey science. We thank Julien Céré who managed the communication between us and the company, and Marine Dupuy who helped us prepare the data used in this study. We thank Alastair J. Wilson for insightful comments on the results. We also thank all members of Pierre-Olivier Montiglio's laboratory who provided useful feedback on earlier versions on this work. This work was supported by an NSERC postgraduate doctoral scholarship (569716-2022).

\newpage

\newpage

# TABLES

@@

# Open the loo table
path <- file.path(here::here("outputs", "02_outputs_model-validation"))
loo_tab <- data.table(readRDS(file.path(path, "A1-3_lootable.rds")))

@@

# Custom theme
my_theme <- function(x, ...) {
   x <- colformat_double(x, big.mark = " ",
                         decimal.mark = ".",
                         digits = 2)
   x <- set_table_properties(x, layout = "autofit")
   x <- border_remove(x)
   std_border <- fp_border(width = 1, color = "black")
   x <- hline_top(x, part = "all", border = std_border)
   x <- hline_bottom(x, part = "all", border = std_border)
}

@@

tab1 <- loo_tab %>%
  select(model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>%
  flextable(col_keys = my_header$col_keys) %>%
  set_header_df(mapping = my_header, key = "col_keys") %>%
  my_theme() %>%
  align(align = "left", part = "all", j = 1) %>%
  align(align = "center", part = "all", j = 2) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5) %>%
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      "* 'elpd' refers to the expected log pointwise density and is the value chosen to select the best model."),
    ref_symbols = " "
  ) %>%

  # Table caption
  set_caption(
    caption = "Table 1. Leave-one-out cross-validation table of the three hierarchical GAMMs relating hunting success to player experience.",
    style = "Table Caption"
  ) %>%

  # Font options
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(size = 12, part = "footer")

tab1

@@

\newpage

@@

# Import the model
fit <- readRDS(
  file.path(path, "01_outputs_models", "B1_DHMLM-no-outlier.rds")
)

@@

  # Align repeated rows on column 1
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top", part = "body") %>%

  # Align text
  align(align = "left", part = "all", j = 1) %>%
  align(align = "left", part = "all", j = 2) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5) %>%

  # Footnote
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "We exponentiated the dispersion parameters (i.e. sigma) which are estimated on a log scale. We back-transformed the hunting success values, estimated on a logit scale, back to a probability scale.",
      as_sup("\nb "),
      "The intercept values on the mean part of the equation for all traits indicate mean behaviour and success at the population level. The intercept values on the dispersion (i.e. sigma) part of the equation for predator speed indicate behavioural specialization at the population level."
    ),
    ref_symbols = " "
  ) %>%

  # Table caption
  set_caption(
    caption = "Table 2. Posterior means and 95% credible intervals of the fixed effects estimated by the MDHGLM of predator speed, prey speed, and predator hunting success.",
    style = "Table Caption"
  ) %>%

  # Font options
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(size = 12, part = "footer")

tab2

@@

\newpage

@@

# Custom theme
my_theme <- function(x, ...) {
  x <- colformat_double(x, big.mark = " ",
                        decimal.mark = ".",
                        digits = 2)
  x <- set_table_properties(x, layout = "autofit")
  x <- border_remove(x)
  std_border <- fp_border(width = 1, color = "black")
  x <- hline_top(x, part = "all", border = std_border)
  x <- hline_bottom(x, part = "all", border = std_border)
}

@@

  # Align text
  align(align = "left", part = "all", j = 1) %>%
  align(align = "left", part = "all", j = 2) %>%
  align(align = "center", part = "all", j = 3) %>%
  align(align = "center", part = "all", j = 4) %>%
  align(align = "center", part = "all", j = 5) %>%

  # Add footnote
  footnote(
    i = 1, j = 1,
    part = "header",
    value = as_paragraph(
      as_sup("a "),
      "We exponentiated the dispersion parameters (i.e. sigma) which are estimated on a log scale. All the reported values are standard deviations.",
      as_sup("\nb "),
      "The intercept values on the mean part of the equation for all traits indicate among individual differences in mean behaviour and success.",
      as_sup("\nc "),
      "The intercept values on the dispersion (i.e. sigma) part of the equation for predator speed indicate among individual differences in behavioural specialization. For prey speed and hunting success, they indicate among individual differences in the variability of prey encounters and variability in hunting success, respectively."
    ),
    ref_symbols = " "
  ) %>%

  # Add table caption
  set_caption(
    caption = "Table 3. Posterior means and 95% credible intervals of the random effects estimated by the MDHGLM of predator speed, prey speed, and predator hunting success.",
    style = "Table Caption"
  ) %>%

  # Font options
  font(fontname = "Times New Roman", part = "all") %>%
  fontsize(size = 10, part = "all") %>%
  fontsize(size = 12, part = "footer")

tab3

@@

\newpage

# FIGURE CAPTIONS

**Figure 1.** Among individual differences in the development of hunting expertise. The predators' hunting success (i.e. the probability of capturing the four prey) is on the y axis, and the predators' cumulative experience (i.e. the number of matches played prior to each observation) is on the x axis. Each fitted curve represents an individual predator. (A) A generalized additive mixed model where we do not control for prey speed (B) A generalized additive mixed model where we control for prey speed.

**Figure 2.** Individual behavioural distribution of the predators' speed drawn from the MDHGLM estimates for individuals that had (A) the greatest increase in specialization and (B) the greatest increase in flexibility. The background (gray) distributions are for when individuals were novice hunters, and the foreground (blue) distributions with solid contour lines are for when they were advanced hunters. Both figure panels are ordered by ascending degree of increase in either specialization or flexibility. We selected the individuals displayed in the figure by first subtracting the estimated standard deviation of all individuals as novices with their standard deviation as advanced hunters, and then selecting those with the greatest difference using the 25% and 75% quantiles. Individuals on panel A have an increase in specialization equal or greater than 0.2 standard deviations, while those on panel B have an increase in flexibility equal or greater than 0.28 standard deviations.

**Figure 3.** Correlation matrices of the MDHGLM relating the mean and intra individual variability (IIV) in speed of the predators, the mean and IIV in speed of the prey they encountered, and mean hunting success. Larger dark blue circles indicate stronger positive correlations (positive values on the color legend), while larger dark red circles indicate stronger negative correlations (negative values on the color legend). A) Correlations when predators were novice hunters. B) Correlations when predators were at an intermediate level of experience. C) Correlations when predators were advanced hunters.

\newpage

# FIGURES

```{r fig1, out.width = "100%", fig.cap = "**Figure 1**"}

knitr::include_graphics(
  here::here("outputs", "04_outputs_figures", "figure1.png")
)
